\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-31 introduction.tex
Updated
Updated the paper

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for correct by construction code generation
and verification of correctness of functional properties of the
developed programs, but leave nonfunctional properties such as timing
behavior as an implementation detail~\cite{boldt07}. Arguably, rightly
so, because physical time cannot be incorporated without information
about the underlying execution platform.  But, one can still reason in
terms of logical time. These languages support discrete logical clock
rather than a discrete physical (real-time) clock. The real-time period
of the discrete logical clock (also referred to as a logical tick) is
not fixed and it is determined by the responsiveness of the program to
the input signals. Unlike a discrete physical clock, which has fixed
real-time period, the period of the logical clock is elastic. Although
the timing model with logical clock has worked well for reactive
synchronous/GALS languages in designing discrete systems, there is a
need to introduce real-time since often, implementation models are in
the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

@@

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
% \item We consider the property of \textit{non
%     maximal-parallelism}. \textit{Non-maximal parallelism:} There are
%   not always sufficient resources for processes to execute, so
%   effectively processes are allocated and scheduled with different
%   optimization criteria -- reducing computation time, reducing power
%   consumption, etc. This allocation and scheduling is an integral part
%   of any language compilation and needs to be considered when
%   introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We shed light upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for correct by construction code generation
and verification of correctness of functional properties of the
developed programs, but leave nonfunctional properties such as timing
behavior as an implementation detail~\cite{boldt07}. Arguably, rightly
so, because physical time cannot be incorporated without information
about the underlying execution platform.  But, one can still reason in
terms of logical time. These languages support discrete logical clock
rather than a discrete physical (real-time) clock. The real-time period
of the discrete logical clock (also referred to as a logical tick) is
not fixed and it is determined by the speed of computation of the
program for different input signals. Unlike a discrete physical clock,
which has fixed real-time period, the period of the logical clock is
elastic. Although, the timing model with logical clock has worked well
for reactive synchronous/GALS languages in designing discrete systems,
there is a need to introduce real-time since often, implementation
models are in the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

@@

\begin{enumerate*}
\item Real-time is specified in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
% \item We consider the property of \textit{non
%     maximal-parallelism}. \textit{Non-maximal parallelism:} There are
%   not always sufficient resources for processes to execute, so
%   effectively processes are allocated and scheduled with different
%   optimization criteria -- reducing computation time, reducing power
%   consumption, etc. This allocation and scheduling is an integral part
%   of any language compilation and needs to be considered when
%   introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We shed light upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

### 2014-01-31 motivation.tex
Updated the paper

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
{ // Clock-domain 1
while(true) {
  trap(T){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(T); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; pause; pause; // pass some time before restarting
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; 
      if(#S == 1) counter++; else counter--;}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

The \texttt{wait\_inbetween (M..N)} statement postpones the program
control flow to move to the proceeding statement for a minimum of $M$
time units and a maximum of $N$ time units, as shown in
Figure~\ref{d:b}.

@@

The SystemJ program is divided into two \hj{mutually asynchronous}
clock-domains, the first clock-domain continuously displays a green
light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
waits for the user to respond. If the user is able to touch the screen
within the specified time of 50.3 ms to 200.3 ms a positive response is
sent to the second clock-domain. All the SystemJ programming constructs
used to implement this system are described in Table~\ref{tab:1} \hj{and
  more details can be found in~\cite{amal10}}. The
\texttt{wait\_inbetween} specification, which models the non exact
postponement between 50.3 ms and 200.3 ms is not supported in the
current language and cannot be compiled with the current SystemJ
compiler. We introduce such \texttt{wait\_inbetween} statements in
SystemJ. Such non exact real-time mechanisms are useful in developing
systems where the real-time postponement is not known or should not be
known a priori.

@@

\begin{figure}[b!]
	\centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{dd:a}Timeout waiting for the input
            signal \texttt{DoorOpened}}
        \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
trap(T){
  abort(DoorOpened){
    wait_atleast(10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{dd:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_atleast} statement. The Y-axis is the
    time elapsed.}
\includegraphics[scale=0.7]{FF2}
\end{SubFloat}
\caption{Programming a timeout with \texttt{wait\_atleast}}
\label{dd}
\end{figure}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. If the signal is not
received then a timeout is generated so that the program can make
progress. Such a timeout programmed in the SystemJ pseudo-code along
with its timing behavior is shown in Figure~\ref{dd}. The SystemJ
program is waiting on an input signal, \texttt{DoorOpened}, from the
environment for 10 seconds and makes progress irrespective of the
reception of this signal after 10 seconds.

@@

\begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exact} statement. The Y-axis is the
    time elapsed.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exact}
  \label{fig:p1}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:p1}. As before upon execution of
\texttt{wait\_exact (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
exactly after 1 ms have passed, as shown in Figure~\ref{p1:b}.

>>>>

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
{ // Clock-domain 1
while(true) {
  trap(T){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(T); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; pause; pause; // pass some time before restarting
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; 
      if(#S == 1) counter++; else counter--;}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_inbetween} statement. The Y-axis is
    the time elapsed. The arrow shows one example of the expiration of
    the \mbox{\texttt{wait\_inbetween (50.3..200.3 ms)}} statement.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

The \texttt{wait\_inbetween (M..N)} statement postpones the program
control flow from moving to the proceeding statement for a minimum of
$M$ time units and a maximum of $N$ time units, as shown in
Figure~\ref{d:b}.

@@

The SystemJ program is divided into two \hj{mutually asynchronous}
clock-domains, the first clock-domain continuously displays a green
light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
waits for the user to respond. If the user is able to touch the screen
within the specified time of at least 50.3 ms and at most 200.3 ms, a
positive response is sent to the second clock-domain. All the SystemJ
programming constructs used to implement this system are described in
Table~\ref{tab:1} \hj{and more details can be found
  in~\cite{amal10}}. The \texttt{wait\_inbetween} specification, which
models the non exact postponement between 50.3 ms and 200.3 ms is not
supported in the current language and cannot be compiled with the
current SystemJ compiler. We introduce such \texttt{wait\_inbetween}
statements in SystemJ. Such non exact real-time mechanisms are useful
for developing systems where the real-time postponement is not known or
should not be known a priori.

@@

\begin{figure}[b!]
	\centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{dd:a}Timeout waiting for the input
            signal \texttt{DoorOpened}}
        \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
trap(T){
  abort(DoorOpened){
    wait_atleast(10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{dd:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_atleast} statement. The Y-axis is the
    time elapsed. The arrow shows one possible expiration of
    \texttt{wait\_atleast (10000 ms)} statement.}
\includegraphics[scale=0.7]{FF2}
\end{SubFloat}
\caption{Programming a timeout with \texttt{wait\_atleast}}
\label{dd}
\end{figure}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. If the signal is not
received, then a timeout is generated so that the program can make
progress. Such a timeout programmed in the SystemJ pseudo-code along
with its timing behavior is shown in Figure~\ref{dd}. The SystemJ
program is waiting on an input signal, \texttt{DoorOpened}, from the
environment for 10 seconds and makes progress irrespective of the
reception of this signal after 10 seconds.

@@

\begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exact} statement. The Y-axis is the
    time elapsed. The arrow shows the \textit{only} possible expiration
    of the \texttt{wait\_exact (1 ms)} statement.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exact}
  \label{fig:p1}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:p1}. As before, upon execution of
\texttt{wait\_exact (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
exactly after 1 ms has passed, as shown in Figure~\ref{p1:b}.
